---
title: "EpiQuant - SARS-Cov epiquant Project (2019-2020)"
author: "Guangzhi, National Microbiology Laboratory "
params:
  clusters:
    input: file
    label: Clusters
    value: clusters/
  cpus:
    input: numeric
    label: CPUs
    value: 1
  gamma:
    input: numeric
    label: γ (geographic)
    value: 1.0
  sigma:
    input: numeric
    label: σ (source)
    value: 0.0
  source:
    input: file
    label: Source Data
    value: source.tsv
  strains:
    input: file
    label: Strain Data
    value: strains.tsv
  tau:
    input: numeric
    label: τ (temporal)
    value: 0.0
output:
  html_document:
    df_print: paged
---


## Introduction 

Hi, thanks for making it this far! Please make sure to read the EpiQuant wiki page, and publication [link] (https://jcm.asm.org/content/55/5/1334) for a better understanding of the methodology. I'm assuming if you've made it this far, you have R and R Studio downloaded. Please read the text under each sections as I have provided additional detail about what the code is doing. I have annotated in the code for additional commments. Let's get started!*


### Section 1: Install the packages needed for the analyses and load them into R 

```{r install-packages, eval=FALSE}

## Install the packages 
install.packages("fossil")
# install.packages("openxlsx")
install.packages("reshape2")
install.packages("plyr")
install.packages("dplyr")
install.packages("tidyr")
install.packages("grid")
install.packages("gridExtra")
install.packages("ggplot2")
install.packages("RColorBrewer")
install.packages("gplots")
install.packages("readr")
install.packages("scatterplot3d")
install.packages("d3heatmap")
install.packages("devtools")
install.packages("github")
install.packages("ggpubr")
install.packages("knitr")
install.packages("ggrepel")
install.packages("stringr")
install.packages("magrittr")
install.packages("tibble")
install.packages("purrr")
install.packages("ggridges")
install.packages("forcats")

```

```{r load-packages}
## Load the packages 
suppressPackageStartupMessages({
library(fossil)
library(openxlsx)
library(reshape2)
library(plyr)
library(dplyr)
library(tidyr)
library(grid)
library(gridExtra)
library(ggplot2)
library(RColorBrewer)
library(gplots)
library(readr)
library(scatterplot3d)
library(d3heatmap)
library(devtools)
library(knitr) 
library(devtools)
# library(ggpubr)
library(ggrepel)
library(stringr)
library(magrittr)
library(tibble)
library(purrr)
library(ggridges)
library(lubridate)
library(forcats)
})

```


### Section 2: Generating the EpiMatrix 

Now, we'll start by loading in a couple different data files: 

1. [**source_pw**](https://www.dropbox.com/s/l1eefag519rpol5/SourcePairwise%2803-Mar-20%29.txt?dl=0) - This file contains the pairwise source similarities generated at the [EpiQuant Webserver] (https://hetmanb.shinyapps.io/EpiQuant/)
2. [**strain_data**](https://www.dropbox.com/s/2mdxskxtma0nox6/Strain_data%2803-Mar-20%29.txt?dl=0) - This file comes from your own data, see below to look at the structure of the file, it needs to be followed precisely for the following scripts to work.  

```{r load-data}
source_pw <- read_tsv(params$source) %>% 
  mutate(Source.Dist = 1- value) %>% 
  select(-value)

strain_data <- read_tsv(params$strains) %>% 
  mutate(Date     = as.Date(paste(Year, Month, Day, sep = "-")),
  Location = paste(Country, Province, City, sep = "_"))

```

```{r calculate-epi-table}
source("helpers/epi-helper.R")

## Coefficients for source, temporal and geographical components of the EpiQuant Model: combination of these should ad  to 1.0
sigma  <- params$sigma #source
tau    <- params$tau   #temporal   
gamma  <- params$gamma #geographical 


# TODO Testing variables: remove
datafile <- strain_data
source_matrix <- source_pw
source_coeff <- sigma
temp_coeff <- tau
geog_coeff <- gamma

## This generates a table of your comparisons based on your epidemiological data (source, time, geographical) with the  assigned weights of s, t and g and then computes the similarity/distance and generates a matrix
epi.table <- EpiTable(strain_data, source_pw, sigma, tau, gamma)
epi.matrix <- EpiMatrix(epi.table)

```

### Section 3: Heatmap for selected clusters

```{r, eval=FALSE}
source("helpers/epi-helper.R")
Strain_Data_EU <- read_csv("Strain_data_USA.csv")
T_3321 <-Strain_Data_EU%>%filter(TP2_cluster_size== 3321)
Matrix_3321 <-epi.matrix[rownames(epi.matrix)%in%T_3321$Strain,colnames(epi.matrix)%in%T_3321 $Strain]
#png("EU_3321_Geo.png")
png("EU_3321_Temp.png")
EpiHeatmap_pdf(Matrix_3321)
dev.off()

T_732 <-Strain_Data_EU%>%filter(TP2_cluster_size== 732)
Matrix_732 <-epi.matrix[rownames(epi.matrix)%in%T_732 $Strain,colnames(epi.matrix)%in%T_732 $Strain]
#png("EU_732_Geo.png")
png("EU_732_Temp.png")
EpiHeatmap_pdf(Matrix_732)
dev.off()

T_593 <-Strain_Data_EU%>%filter(TP2_cluster_size== 593)
Matrix_593 <-epi.matrix[rownames(epi.matrix)%in%T_593 $Strain,colnames(epi.matrix)%in%T_593 $Strain]
#png("EU_593_Geo.png")
png("EU_593_Temp.png")
EpiHeatmap_pdf(Matrix_593)
dev.off()

T_525 <-Strain_Data_EU%>%filter(TP2_cluster_size== 525)
Matrix_525 <-epi.matrix[rownames(epi.matrix)%in%T_525 $Strain,colnames(epi.matrix)%in%T_525 $Strain]
png("EU_525_Geo.png")
#png("EU_525_Temp.png")
EpiHeatmap_pdf(Matrix_525)
dev.off()

T_264 <-Strain_Data_EU%>%filter(TP2_cluster_size== 264)
Matrix_264 <-epi.matrix[rownames(epi.matrix)%in%T_264$Strain,colnames(epi.matrix)%in%T_264 $Strain]
png("EU_264_Geo.png")
#png("EU_264_Temp.png")
EpiHeatmap_pdf(Matrix_264)
dev.off()
```

            